US 2013-2015 Fatality Analysis Reporting System (FARS) exploratory data analysis
========================================================
##### by Shiu-Tang Li
###1. Brief introduction, and some remarks to this interactive EDA
In this report, A few interactive maps / pie charts are created to help the readers gain intuition on FARS dataset, including the visualizations of a few major factors (drunken driving, non-motor vehicle occupant deaths, weather, number of persons died in an accident).

**The data set**

FARS 2013-2015 dataset contains 50-52 features (differed by years), and contains 92,424 observations (30202, 30056, 32166 obs for year 2013, 2014, 2015).

The FARS data set can be downloaded from ftp://ftp.nhtsa.dot.gov/fars/, and the full description of the FARS dataset can be downloaded from 
https://crashstats.nhtsa.dot.gov/Api/Public/Publication/812315 .

In this report, all data with wrong longitude or latitude (say latitude = 999.99) are removed (722 obs). 

**Additional Information as you hover the cursor over the map..**

Format: Month/Day/Year(Weekday), 24 hour time format, number of persons killed, first harmful event (major reason that causes death) 

**Create a customized EDA report!**

You could create an analysis report for any state you like, by specifying the state abbreviations (All 50 states + Washington D.C., not including Virgin Islands and Puerto Rico because there're no data) in the .rmd file. The default is "UT" (Utah).  

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
# Choose a state to analyze (Use Capitalized caracters!)
state_abbr <- "UT"

# Virgin Islands contains 0 obs in year 2013-2015
# Puerto Rico contains 0 obs in year 2013-2015
```

```{r packages}
#### load all libraries, functions, and data
library(png)       # show images on rmd
library(grid)      # show images on rmd
library(dplyr)
library(leaflet)   # for interactive maps
library(plotly)    # for pie chart
library(foreign)   # to load .dbf file

# Some settings to get interactive map working in .html 
addTiles = function(map, 
                    urlTemplate = "http://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png",
                    attribution = NULL, 
                    layerId = NULL, 
                    group = NULL, 
                    options = tileOptions()){
options$attribution = attribution
if (missing(urlTemplate) && is.null(options$attribution))
options$attribution = paste("OpenStreetMap",
"contributors, CC-BY-SA")
invokeMethod(map, getMapData(map), "addTiles", urlTemplate,
layerId, group, options)
}


# load functions / decoders
zero_or_not <- function(value){
  if (value == 0) {
    return("No")
  }
  else{
    return("Yes")
  }
}

missing_values_replacement <- function(value){
  if (value == '99') {
    return("--")
  }
  else{
    return(value)
  }
}

weather_decoding <- function(value){
  if (value == 1) {
    return("Clear")
  }
  else if (value == 0){
    return("No Additional Atmospheric Conditions")
  }
  else if (value == 2){
    return("Rain")
  }
  else if (value == 3){
    return("Sleet, Hail")
  }
  else if (value == 4){
    return("Snow")
  }
  else if (value == 5){
    return("Fog, Smog, Smoke")
  }
  else if (value == 6){
    return("Severe Crosswinds")
  }
  else if (value == 7){
    return("Blowing Sand, Soil, Dirt")
  }
  else if (value == 8){
    return("Other")
  }
  else if (value == 10){
    return("Cloudy")
  }
  else if (value == 11){
    return("Blowing Snow")
  }
  else if (value == 12){
    return("Freezing Rain or Drizzle")
  }
  else{
    return("Unknown")
  }
}

weekday_decoding <- function(value){
  if (value == 1) {
    return("(Sun)")
  }
  else if (value == 2){
    return("(Mon)")
  }
  else if (value == 3){
    return("(Tue)")
  }
  else if (value == 4){
    return("(Wed)")
  }
  else if (value == 5){
    return("(Thu)")
  }
  else if (value == 6){
    return("(Fri)")
  }
  else if (value == 7){
    return("(Sat)")
  }
  else{
    return("--")
  }
}

first_harmful_event_decoding <- function(value){
  if (value == 1) {
    return("Rollover/Overturn")
  }
  else if (value == 2){
    return("Fire/Explosion")
  }
  else if (value == 3){
    return("Immersion")
  }
  else if (value == 4){
    return("Gas Inhalation")
  }
  else if (value == 5){
    return("Fell/Jumped from Vehicle")
  }
  else if (value == 6){
    return("Injured in Vehicle (Non-Collision)")
  }
  else if (value == 7){
    return("Other Non-Collision")
  }
  else if (value == 8){
    return("Pedestrian")
  }
  else if (value == 9){
    return("Pedalcyclist")
  }
  else if (value == 10){
    return("Railway Vehicle")
  }
  else if (value == 11){
    return("Live Animal")
  }
  else if (value == 12){
    return("Motor Vehicle in Transport")
  }
  else if (value == 14){
    return("Parked Motor Vehicle (Not In Transport)")
  }
  else if (value == 15){
    return("Non-Motorist on Personal Conveyance")
  }
  else if (value == 16){
    return("Thrown or Falling Object")
  }
  else if (value == 17){
    return("Boulder")
  }
  else if (value == 18){
    return("Other Object (Not Fixed)")
  }
  else if (value == 19){
    return("Building")
  }
  else if (value == 20){
    return("Impact attenuator")
  }
  else if (value == 21){
    return("Bridge Pier or Support")
  }
  else if (value == 23){
    return("Bridge Rail (Includes Parapet)")
  }
  else if (value == 24){
    return("Guardrail Face")
  }
  else if (value == 25){
    return("Concrete Traffic Barrier")
  }
  else if (value == 26){
    return("Other Traffic Barrier")
  }
  else if (value == 30){
    return("Utility Pole/Light Support")
  }
  else if (value == 31) {
    return("Other Post/Pole/Support")
  }
  else if (value == 32){
    return("Culvert")
  }
  else if (value == 33){
    return("Curb")
  }
  else if (value == 34){
    return("Ditch")
  }
  else if (value == 35){
    return("Embankment")
  }
  else if (value == 38){
    return("Fence")
  }
  else if (value == 39){
    return("Wall")
  }
  else if (value == 40){
    return("Fire Hydrant")
  }
  else if (value == 41){
    return("Shrubbery")
  }
  else if (value == 42){
    return("Tree (Standing Only)")
  }
  else if (value == 43){
    return("Other Fixed Object")
  }
  else if (value == 44){
    return("Pavement Surface Irregularity")
  }
  else if (value == 45){
    return("Working Motor Vehicle")
  }
  else if (value == 46){
    return("Traffic Signal Support")
  }
  else if (value == 48){
    return("Snow Bank")
  }
  else if (value == 49){
    return("Ridden Animal or Animal-Drawn Conveyance")
  }
  else if (value == 50){
    return("Bridge Overhead Structure")
  }
  else if (value == 51){
    return("Jackknife")
  }
  else if (value == 52){
    return("Guardrail end")
  }
  else if (value == 53){
    return("Mail Box")
  }
  else if (value == 54){
    return("Motor Vehicle In-Transport Collisions")
  }
  else if (value == 55){
    return("Motor Vehicle in Motion Outside the Trafficway")
  }
  else if (value == 57){
    return("Cable Barrier")
  }
  else if (value == 58){
    return("Ground")
  }
  else if (value == 59){
    return("Traffic Sign Support")
  }
  else if (value == 72){
    return("Cargo/Equipment Loss or Shift")
  }
  else if (value == 73){
    return("Object Fell From Motor Vehicle In-Transport ")
  }
  else{
    return("Unknown")
  }
}

## State switch
switch(state_abbr, 
UT={
state_number<- 49 
state_name  <- "Utah"
},
AL={
state_number<- 1 
state_name  <- "Alabama"   
},
AK={
state_number<- 2 
state_name  <- "Alaska"   
},
AZ={
state_number<- 4 
state_name  <- "Arizona"
},
AR={
state_number<- 5 
state_name  <- "Arkansas"   
},
CA={
state_number<- 6 
state_name  <- "California"   
},
CO={
state_number<- 8 
state_name  <- "Colorado"
},
CT={
state_number<- 9 
state_name  <- "Connecticut"   
},
DE={
state_number<- 10 
state_name  <- "Delaware"   
},
DC={
state_number<- 11 
state_name  <- "District of Columbia"
},
FL={
state_number<- 12
state_name  <- "Florida"   
},
GA={
state_number<- 13 
state_name  <- "Georgia"   
},
HI={
state_number<- 15 
state_name  <- "Hawaii"
},
ID={
state_number<- 16 
state_name  <- "Idaho"   
},
IL={
state_number<- 17 
state_name  <- "Illinois"   
},
IN={
state_number<- 18
state_name  <- "Indiana"
},
IA={
state_number<- 19
state_name  <- "Iowa"   
},
KS={
state_number<- 20
state_name  <- "Kansas"   
},
KY={
state_number<- 21
state_name  <- "Kentucky"
},
LA={
state_number<- 22 
state_name  <- "Louisiana"   
},
ME={
state_number<- 23 
state_name  <- "Maine"   
},
MD={
state_number<- 24
state_name  <- "Maryland"
},
MA={
state_number<- 25
state_name  <- "Massachusetts"   
},
MI={
state_number<- 26 
state_name  <- "Michigan"   
},
MN={
state_number<- 27
state_name  <- "Minnesota"
},
MS={
state_number<- 28
state_name  <- "Mississippi"   
},
MO={
state_number<- 29
state_name  <- "Missouri"   
},
MT={
state_number<- 30 
state_name  <- "Montana"
},
NE={
state_number<- 31
state_name  <- "Nebraska"   
},
NV={
state_number<- 32
state_name  <- "Nevada"   
},
NH={
state_number<- 33 
state_name  <- "New Hampshire"
},
NJ={
state_number<- 34 
state_name  <- "New Jersey"   
},
NM={
state_number<- 35 
state_name  <- "New Mexico"   
},
NY={
state_number<- 36
state_name  <- "New York"
},
NC={
state_number<- 37
state_name  <- "North Carolina"   
},
ND={
state_number<- 38
state_name  <- "North Dakota"   
},
OH={
state_number<- 39 
state_name  <- "Ohio"
},
OK={
state_number<- 40 
state_name  <- "Oklahoma"   
},
OR={
state_number<- 41 
state_name  <- "Oregon"   
},
PA={
state_number<- 42
state_name  <- "Pennsylvania"
},
#PR={   # no data in 2013-2015
#state_number<- 43
#state_name  <- "Puerto Rico"   
#},
RI={
state_number<- 44
state_name  <- "Rhode Island"   
},
SC={
state_number<- 45
state_name  <- "South Carolina"
},
SD={
state_number<- 46 
state_name  <- "South Dakota"   
},
TN={
state_number<- 47 
state_name  <- "Tennessee"   
},
TX={
state_number<- 48
state_name  <- "Texas"
},
VT={
state_number<- 50
state_name  <- "Vermont"   
},
#VI={       # no data in 2013-2015
#state_number<- 52 
#state_name  <- "Virgin Islands"   
#},
VA={
state_number<- 51
state_name  <- "Virginia"
},
WA={
state_number<- 53
state_name  <- "Washington"
},
WV={
state_number<- 54
state_name  <- "West Virginia"   
},
WI={
state_number<- 55
state_name  <- "Wisconsin"   
},
WY={
state_number<- 56
state_name  <- "Wyoming"
}
)

## load data / combine data / preprocessing

df <- read.csv("FARS2015NationalCSV/accident.csv")
df <- df[,c("STATE", "PEDS", "DAY", "MONTH",
            "YEAR", "DAY_WEEK", "HOUR", "MINUTE",
            "LATITUDE", "LONGITUD", "HARM_EV",
            "WEATHER", "FATALS", "DRUNK_DR")]
df <- df[(df$LATITUDE <= 90) &
         (df$LONGITUD <= 180), ]

df_2014 <- read.dbf("FARS2014NationalDBF/accident.dbf")
df_2014 <- df_2014[,c("STATE", 
            "PEDS", "DAY","MONTH",
            "YEAR", "DAY_WEEK", "HOUR", "MINUTE",
            "LATITUDE", "LONGITUD", "HARM_EV",
            "WEATHER", "FATALS", "DRUNK_DR")]
df_2014 <- df_2014[(df_2014$LATITUDE <= 90) &
                   (df_2014$LONGITUD <= 180), ]

df_2013 <- read.dbf("FARS2013NationalDBF/accident.dbf")
df_2013 <- df_2013[,c("STATE", 
            "PEDS", "DAY","MONTH",
            "YEAR", "DAY_WEEK", "HOUR", "MINUTE",
            "LATITUDE", "LONGITUD", "HARM_EV",
            "WEATHER", "FATALS", "DRUNK_DR")]
df_2013 <- df_2013[(df_2013$LATITUDE <= 90) &
                   (df_2013$LONGITUD <= 180), ]


df <- rbind(df, df_2014)
df <- rbind(df, df_2013)

df_state <- df[df$STATE == state_number, ]  

Hour  <- sapply(df_state$HOUR, toString)
Hour  <- sapply(Hour, missing_values_replacement)
Min   <- sapply(df_state$MINUTE + 100, toString)
Min   <- sapply(Min, substr, 2, 3)
Min   <- sapply(Min, missing_values_replacement)
Weekday <- sapply(df_state$DAY_WEEK, weekday_decoding)
Month <- sapply(df_state$MONTH, toString)
Day   <- sapply(df_state$DAY, toString)
Year  <- sapply(df_state$YEAR - 2000, toString)

#Weather<- sapply(df_state$WEATHER, weather_decoding)

Acc_Time  <- paste(Month, Day, sep="/")
Acc_Time  <- paste(Acc_Time, Year, sep="/")
Acc_Time  <- paste(Acc_Time, Weekday, sep="")
Acc_Time  <- paste(Acc_Time, Hour, sep=", ")
Acc_Time  <- paste(Acc_Time, Min, sep=":")
df_state$info <- Acc_Time

number_killed <- sapply(df_state$FATALS, toString)
df_state$info <- paste(df_state$info, 
                       number_killed, sep =", ")
df_state$info <- paste(df_state$info, 
                       "killed. ", sep =" ")
first_harmful_event <- sapply(df_state$HARM_EV,
                              first_harmful_event_decoding)

df_state$info <- paste(df_state$info, 
                       first_harmful_event, ".",
                       sep ="")

# rm(list=ls())
```

```{r fig.width=5, fig.height=1}
img <- readPNG("set_state.png")
grid.raster(img)
``` 

###2. Drunken driving

```{r}
df2 <- data.frame(df_state$LATITUDE, 
                  df_state$LONGITUD, 
                  df_state$info, 
                  df_state$DRUNK_DR)
colnames(df2) <- c("lat", "long", "info", "drunk")
df2$drunk <- sapply(df2$drunk, zero_or_not)

df3 <- data.frame(df$DRUNK_DR)
colnames(df3) <- c("drunk")
df3$drunk <- sapply(df3$drunk, zero_or_not)

p1<- plot_ly(data = count(df2, drunk), 
             type = 'pie',
             textposition = 'inside',
             textinfo = 'label+percent',
             insidetextfont = list(color = '#FFFFFF'),
             labels = ~drunk, 
             values = ~n,
             hoverinfo = "text",
             text = ~n,
             domain = list(x = c(0, 0.47), 
                           y = c(0, 0.94)),
             marker = list(line = 
                           list(color = '#FFFFFF', 
                                width = 1))) 

p2<- plot_ly(data = count(df3, drunk), 
             type = 'pie',
             textposition = 'inside',
             textinfo = 'label+percent',
             insidetextfont = list(color = '#FFFFFF'),
             labels = ~drunk, 
             values = ~n,
             hoverinfo = "text",
             text = ~n, 
             domain = list(x = c(0.53, 1), 
                           y = c(0, 0.94)),
             marker = list(line = 
                           list(color = '#FFFFFF', 
                                width = 1))) 

title_string <- paste("Drunken Driving in 2013-2015 traffic fatalities: ", state_name, " vs Nation", sep="")

subplot(p1, p2, nrows = 1) %>% 
layout(title = title_string, 
       showlegend = F,
       margin = list(l = 50,r = 50,b = 10,t = 80,
                     pad = 4),
       xaxis = list(showgrid = FALSE, 
                    zeroline = FALSE, 
                    showticklabels = FALSE),
       yaxis = list(showgrid = FALSE, 
                    zeroline = FALSE, 
                    showticklabels = FALSE),
       annotations = list(list(x = 0.22, 
                               y = 0.95, 
                               text = state_abbr, 
                               showarrow = F,
                               xref='paper', 
                               yref='paper'),
                          list(x = 0.81, 
                               y = 0.95, 
                               text = "Nation", 
                               showarrow = F,
                               xref='paper', 
                               yref='paper')))
```  

**2013-2015 `r state_name` traffic fatalities by drunken driving**
```{r}
palette <- colorFactor(c("black", "red"), domain = c("No", "Yes"))

leaflet(data = df2) %>% 
addTiles() %>% 
addCircleMarkers(~long, ~lat
                ,label  = ~as.character(info)
                ,color  = ~palette(drunk)
                ,radius = 6
                ,stroke = FALSE
                ,fillOpacity = 0.7 
                )%>% 
addLegend("bottomright", 
          pal = palette,
          values = ~drunk,
          title  = "Drunken driving",
          opacity = 1
          )
```

###3. Non-Motor Vehicle Occupant killed in an accident

Non-motor vehicle occupants refer to people NOT in motor vehicles, e.g. pedestrians, cyclists.

```{r}
df2 <- data.frame(df_state$LATITUDE, 
                  df_state$LONGITUD, 
                  df_state$info, 
                  df_state$PEDS)
colnames(df2) <- c("lat", "long", "info", "peds")
df2$peds <- sapply(df2$peds, zero_or_not)

df3 <- data.frame(df$PEDS)
colnames(df3) <- c("peds")
df3$peds <- sapply(df3$peds, zero_or_not)

p1<- plot_ly(data = count(df2, peds), 
             type = 'pie',
             textposition = 'inside',
             textinfo = 'label+percent',
             insidetextfont = list(color = '#FFFFFF'),
             labels = ~peds, 
             values = ~n,
             hoverinfo = "text",
             text = ~n,
             domain = list(x = c(0, 0.47), 
                           y = c(0, 0.94)),
             marker = list(line = 
                           list(color = '#FFFFFF', 
                                width = 1))) 

p2<- plot_ly(data = count(df3, peds), 
             type = 'pie',
             textposition = 'inside',
             textinfo = 'label+percent',
             insidetextfont = list(color = '#FFFFFF'),
             labels = ~peds, 
             values = ~n,
             hoverinfo = "text",
             text = ~n, 
             domain = list(x = c(0.53, 1), 
                           y = c(0, 0.94)),
             marker = list(line = 
                           list(color = '#FFFFFF', 
                                width = 1))) 
title_string <- paste("Are Non-Motor Vehicle Occupants killed in 2013-2015 traffic fatalities?<br>", state_name, " vs Nation", sep="")

subplot(p1, p2, nrows = 1) %>% 
layout(title = title_string, 
       showlegend = F,
       margin = list(l = 50,r = 50,b = 10,t = 80,
                     pad = 4),
       xaxis = list(showgrid = FALSE, 
                    zeroline = FALSE, 
                    showticklabels = FALSE),
       yaxis = list(showgrid = FALSE, 
                    zeroline = FALSE, 
                    showticklabels = FALSE),
       annotations = list(list(x = 0.22, 
                               y = 0.95, 
                               text = state_abbr, 
                               showarrow = F,
                               xref='paper', 
                               yref='paper'),
                          list(x = 0.81, 
                               y = 0.95, 
                               text = "Nation", 
                               showarrow = F,
                               xref='paper', 
                               yref='paper')))
```  

**2013-2015 `r state_name` traffic fatalities by non-motor vehicle occupant deaths**
```{r}
palette <- colorFactor(c("black", '#00c800'), domain = c("No", "Yes"))

leaflet(data = df2) %>% 
addTiles() %>% 
addCircleMarkers(~long, ~lat
                ,label  = ~as.character(info)
                ,color  = ~palette(peds)
                ,radius = 6
                ,stroke = FALSE
                ,fillOpacity = 0.7 
                )%>% 
addLegend("bottomright", 
          pal = palette,
          values = ~peds,
          title  = "Non-motor vehicle <br> occupants killed",
          opacity = 1
          )
```


###4. Weather conditions

```{r}
df2 <- data.frame(df_state$LATITUDE, 
                  df_state$LONGITUD, 
                  df_state$info, 
                  df_state$WEATHER)
colnames(df2) <- c("lat", "long", "info", "weather")
df2$weather <- sapply(df2$weather, weather_decoding)

df3 <- data.frame(df$WEATHER)
colnames(df3) <- c("weather")
df3$weather <- sapply(df3$weather, weather_decoding)


p1<- plot_ly(data = count(df2, weather), 
             type = 'pie',
             textposition = 'inside',
             textinfo = 'label+percent',
             insidetextfont = list(color = '#FFFFFF'),
             labels = ~weather, 
             values = ~n,
             name = "Weather",
             domain = list(x = c(0, 0.47), 
                           y = c(0, 0.94)),
             marker = list(line = 
                           list(color = '#FFFFFF', 
                                width = 1))) 

p2<- plot_ly(data = count(df3, weather), 
             type = 'pie',
             textposition = 'inside',
             textinfo = 'label+percent',
             insidetextfont = list(color = '#FFFFFF'),
             labels = ~weather, 
             values = ~n,
             name = "Weather",
             domain = list(x = c(0.53, 1), 
                           y = c(0, 0.94)),
             marker = list(line = 
                           list(color = '#FFFFFF', 
                                width = 1))) 

title_string <- paste("weather conditions in 2013-2015 traffic fatalities:<br>", state_name, " vs Nation", sep="")

subplot(p1, p2, nrows = 1) %>% 
layout(title = title_string, 
       showlegend = F,
       margin = list(l = 50,r = 50,b = 10,t = 80,
                     pad = 4),
       xaxis = list(showgrid = FALSE, 
                    zeroline = FALSE, 
                    showticklabels = FALSE),
       yaxis = list(showgrid = FALSE, 
                    zeroline = FALSE, 
                    showticklabels = FALSE),
       annotations = list(list(x = 0.22, 
                               y = 0.95, 
                               text = state_abbr, 
                               showarrow = F,
                               xref='paper', 
                               yref='paper'),
                          list(x = 0.81, 
                               y = 0.95, 
                               text = "Nation", 
                               showarrow = F,
                               xref='paper', 
                               yref='paper')))
```  

**2013-2015 `r state_name` traffic fatalities by weather conditions**
```{r}
#pal <- colorFactor(
#  palette = "Spectral",
#  domain  = df2$weather
#)

palette <- colorFactor(c("#994C00", 
                         "#00FFFF",
                         "yellow", 
                         "orange",
                         "#CC00CC",
                         "#006633",
                         "#E0E0E0",
                         "#606060",
                         '#00c800', 
                         "#FF3333",
                         '#00A8A8',
                         '#0000FF', 
                         "black"), 
                       
              domain = c("Blowing Sand, Soil, Dirt",
                         "Blowing Snow",
                         "Clear",
                         "Cloudy",
                         "Fog, Smog, Smoke",
                         "Freezing Rain or Drizzle",
                         "No Additional Atmospheric Conditions",
                         "Other",
                         "Rain",
                         "Severe Crosswinds",
                         "Sleet, Hail",
                         "Snow",
                         "Unknown"))



leaflet(data = df2) %>% 
addTiles() %>% 
addCircleMarkers(~long, ~lat
                ,label  = ~as.character(info)
                ,color  = ~palette(weather)
                ,radius = 6
                ,stroke = FALSE
                ,fillOpacity = 0.8 
                )%>% 
addLegend("bottomright", 
          pal = palette,
          values = ~weather,
          title  = "Weather conditions",
          opacity = 1
          )
```

###5. Number of fatally injured persons in the crash

```{r}
df2 <- data.frame(df_state$LATITUDE, 
                  df_state$LONGITUD, 
                  df_state$info, 
                  df_state$FATALS)
colnames(df2) <- c("lat", "long", "info", "fatals")

df3 <- data.frame(df$FATALS)
colnames(df3) <- c("fatals")

```
**Year 2013-2015: number of fatally injured persons in the crash in `r state_name` (Number of deaths / counts)** 
```{r}
table(df2$fatals)
```
**Year 2013-2015: number of fatally injured persons in the crash nationally (Number of deaths / counts)**
```{r}
table(df3$fatals)
```

**2013-2015 `r state_name` traffic fatalities by number of deaths in one accident (the larger the bubble size the larger the number)**
```{r}
#p1 <- plot_ly(count(df2, fatals), 
#             x = ~fatals, 
#             y = ~n, 
#             type = 'bar')

leaflet(data = df2) %>% 
addTiles() %>% 
addCircleMarkers(~long, ~lat
                ,label  = ~as.character(info)
                ,color = 'black',
                ,radius = ~as.numeric(fatals)*1.5+2
                ,stroke = FALSE
                ,fillOpacity = 0.7 
                )
```


```{r}
# remove outliers
### DE   2 in wrong state

```
